---
title: "Chocolate Chip Cookies"
execute:
  error: true
author: "Luke Nieland"
output: html_document
---

## Reading In the Data

First, read in the CSV data of cookie ingredients.
Make sure that your end-result data has appropriate types for each column - these should match the types provided in the documentation in the README.md file.

```{r}
data <- read.csv('choc_chip_cookie_ingredients.csv')

column_types <- sapply(data, class)

column_types #Unit is not numeric like in ReadMe, however the values within the column some need to be text, so convert to factor to account for that

data$Unit <- as.factor(data$Unit) #convert Unit to factor

column_types_fixed <- sapply(data, class)

column_types_fixed

```

```{python}

```


## Exploratory Data Analysis

Exploratory data analysis is the process of getting familiar with your dataset. To get started, [this blog post](https://www.mrdbourke.com/a-gentle-introduction-to-exploratory-data-analysis/) provides a nice checklist to get you thinking:

> 1.  What question(s) are you trying to solve (or prove wrong)?
> 2.  What kind of data do you have and how do you treat different types?
> 3.  What's missing from the data and how do you deal with it?
> 4.  Where are the outliers and why should you care about them?
> 5.  How can you add, change or remove features to get more out of your data?

### Generating Questions

Generate at least 5 questions you might explore using this database of cookie ingredients.

1. Are there any missing or NA rows/columns/data points? Do they need to be deleted or have data imputed?
2. What ingredients are used in the most and least recipes?
3. What is the average amount of cookies made by a recipe?
4. How many recipes of each rating do we have?
5. Are there any outliers? Does it effect rating.


### Skimming the Data

One thing we often want to do during EDA is to examine the quality of the data - are there missing values? What quirks might exist in the dataset?

The `skimr` package in R, and the similar `skimpy` package in python (which has a much better name, in my opinion), can help provide visual summaries of the data. 

Install both packages, and read the package documentation ([R](https://cran.r-project.org/web/packages/skimr/vignettes/skimr.html), [Python](https://pypi.org/project/skimpy/)).

[Part 1] Use each package and generate summaries of your data that require the use of at least some non-default options in each package's `skim` function.


```{r}
library(skimr)

my_skim <- skim_with(numeric = sfl(hist = NULL, iqr = ~ IQR(., na.rm = TRUE), p25 = NULL, p75 = NULL)) #removed historgram and p25, p75 and added IQR

my_skim(data)

```

```{python}

```

[Part 2] Write 1-2 sentences about what you can tell from each summary display you generate. Did you discover anything new about the data?

The character data type columns are fully complete without any nulls, however the rating column in the numeric columns is only about 49% complete. As well, the IQR for quantity is the largest, meaning many of these recipes produce differing amounts of cookies.

### Generating Tables

Another useful technique for exploratory data analysis is to generate summary tables. 
You may want to use the `dplyr` package in R (`group_by` or `count` functions), as well as the `groupby` and `count` methods in Pandas. [Python example](https://sparkbyexamples.com/pandas/pandas-groupby-count-examples/), [R example](https://dplyr.tidyverse.org/reference/count.html)

[Part 1] Using R and Python, generate a table that shows what **proportion** of recipes contain each type of ingredient, for the most common 20 ingredients.

```{r}
library(tidyverse)

total_recipes <- nrow(distinct(data, Recipe_Index))

top20_props <- data %>% #this pipe thing I learned about while looking up stuff about dplyr, pretty nice tool
  distinct(Recipe_Index, Ingredient) %>% #each unique ingredient pair
  count(Ingredient, name = "n_recipes") %>% #how many recipes each ingredient is in
  mutate(proportion = n_recipes / total_recipes) %>% #proportion of number of recipes in to total recipes
  arrange(desc(proportion)) %>%
  slice_head(n = 22) #get top 20 (22 to account for ties in last place)

top20_props
```


[Part 2] Print out a character string that lists all of the ingredients that do not appear in at least 20 recipes.

```{r}
library(tidyverse)

not_top20_props <- data %>% #this pipe thing I learned about while looking up stuff about dplyr, pretty nice tool
  distinct(Recipe_Index, Ingredient) %>% #each unique ingredient pair
  count(Ingredient, name = "n_recipes") %>% #how many recipes each ingredient is in
  filter(n_recipes < 20) %>% #gets ingredients in less than 20 recipes
  pull(Ingredient) #pulls out the names of those ingredients.

paste(not_top20_props, collapse = " ") #prints character string
```


(Delete this note, but you can include data values inline in markdown text by using backticks, at least in R. For instance, here is R's built in value for pi: `r pi`. Unfortunately, this doesn't work in python using the knitr markdown engine, but you can print the list out in python anyways using a code chunk.)

### Visualization

Using whatever plotting system you are comfortable with in R or python, see if you can create a couple of useful exploratory data visualizations which address one of the questions you wrote above - or another question which you've come up with as you've worked on this assignment.

[Part 1] Create at least one plot (it doesn't have to be pretty) that showcases an interesting facet of the data.
```{r}
ingred <- table(data$Ingredient)

barplot(ingred,
        main = 'Bar Chart of Ingredient Counts',
        xlab = 'Ingredient',
        ylab = '# times Used')

```


[Part 2] Write 2-3 sentences about what you can learn from that plot and what directions you might want to investigate from here.

From the plot, I can answer my question about which ingredients are used the most. While we answered this earlier, getting a visual helps me understand how the most common ingredients are used much more frequently than the lesser used ingredients. A further investigation could be the lowest ingredients used and visualize those to compare the more "frequent" infrequent ingredients and the truly rarely used ingredients.
